Agile Monitoring of Evolving Web Services: For the Little Man

### 2018-04-06 evolution.tex
...

  The Stack Overflow Developer Survey from 2018 revealed the fact that 88.4\% of professional developers use git for version control. 
  Given that versioning using git is the main way in which the source code of services evolves and is deployed, it is natural to monitor their progress also across versions.

  Version control in \tool can be supported in two ways,
  the developer explicitly states the current version, 
  or the current version is automatically detected based
  on some version control system. 
  The second option is the most interesting one.
  For the systems deployed using Git, an extra configuration line enables \tool to automatically\footnote{Alternatively, the maintainer can add version identifiers manually for the web application through a configuration file if the system does not use git.} associate API traffic with the currently deployed version based on information in the .git folder \footnote{\url{https://git-scm.com/}}:

@@

  \Fref{fig:mv-util} shows the \perspective{Multi-Version API Utilization} perspective which presents the utilization of the tracked endpoints across versions. Because the different versions might be deployed for very different periods of time, at the intersection of an endpoint and a version the chart does not plot the absolute utilization of that endpoint in that version but rather the percentage of all the API calls that go to that endpoint. Otherwise a version that is deployed for many weeks would make all those deployed for several days invisible.

@@

  With this view, several patterns are visible:
  \begin{itemize}

    \item \epFeedItems is discontinued a little bit after \epTopArticles is being introduced

>>>>

  The Stack Overflow Developer Survey from 2018 revealed the fact that 88.4\% of professional developers use Git. 
  Given that versioning using git is the main way in which the source code of services evolves and is deployed, it is natural to monitor their progress also across versions.

  Version control in a dashboard can be supported in two ways: 
  (1) the developer explicitly states the current version, or 
  (2) the current version is automatically detected based on some version control system.

  For \tool to start using Git, an extra configuration line enables it to automatically\footnote{Alternatively, the maintainer can add version identifiers manually for the web application through a configuration file if the system does not use Git.} associate API traffic with the currently deployed version based on information in the .git folder \footnote{\url{https://git-scm.com/}}:

@@

  \Fref{fig:mv-util} shows the \perspective{Multi-Version API Utilization} perspective which presents the utilization of the tracked endpoints across versions. The view is a matrix, in which, the intensity of the color at the intersection of an endpoint line and a version column is proportional to the percentage of requests served by that endpoint in that particular version\footnote{The chart does not plot the absolute utilization of that endpoint in that version but rather the percentage of all the API calls that go to that endpoint. Otherwise a version that is deployed for many weeks would make all those deployed for a few days invisible}.

@@

  With this view, several patterns could be visible on such a graph:
  \begin{itemize}

    \item Endpoints being deprecateds. In the case study \epFeedItems stops being used somewhere after the half of the presented timeline.

    \item Endpoints being introduced. In the case study, \epTopArticles appears soon after the half o the presented timeline.

    \item Endpoint renames. Although not visible in the figure, the candidates for rename would be visible as one would discontinued at the same time when the other would appear.

### 2018-04-06 grouping.tex
...

It is often the case that a service has diferent types of users: the paying and the free-for-evaluation ones, or the in-house and those in a different organization.

In some other cases, it might be important for the maintainer of the API to understand the performance on a per-user basis, especially for situations where the system response time is a function of some individual user load\footnote{E.g. in GMail some users have a few dozen emails while other have tens of thousands: this difference in user loads will eventually induce a difference in the response times for different users}.

To support narrowing down the analysis to groups of users, or even individual users, the \tool must log together with every request grouping information for that request. The simplest way of achieving this is to take advantage of the architecture of Flask applications in which a global \code{flask.request} is used to retrieve session information which can in turn is normally used to identify the user sending the request. From the user one can obtain the group.

The following code snippet shows how to configure the dashboard to enable user-by-user analysis\footnote{	If we wanted to group by the user group for example, the code would change slightly by replacing ``user.id'' with ``user.group.id''}:

@@

% \niceseparator

In Zeeguu, \epFeedItems retrieves a list of recommended articles for a given user. Cf. \Fref{fig:ep} it is the endpoint with the slowest response time and highest variability. The reason for this is that a user can be subscribed to anything from one to three dozen article sources and for each of the sources the system must compute the personalized difficulty of each article at every request.

@@

A \perspective{Per-User Performance} perspective should show the different response times for different users. Figure \ref{fig:tpu} presents a subset of the corresponding view in the \tool. The figure shows that the response times for this endpoint can vary considerably for different users with some extreme cases where a user has to wait a full minute until their recommended articles are shown\footnote{After seeing this perspective, the maintainer refactored the architecture of the system to move part the difficulty computation out of the interactive loop}.

>>>>

It is sometimes the case that the utilization and performance of an API must be understood not only per endpoint, but also by grouping the requests by another criterion. Two such examples are: 
\begin{itemize}
	\item When a service has diferent types of users (e.g. the paying vs. the free-for-evaluation ones, or the in-house vs. those in a different organization) the maintainer must understand the performance for the different groups
	\item When the load mix o the users can vary dramatically and the system response time is a function of the individual user load\footnote{E.g. in GMail some users have a few dozen emails while other have tens of thousands: this difference in user loads will eventually induce a difference in the response times for different users} the maintainer must understand the variation of performance with the users (and implicitly the user load). s
\end{itemize}

To support narrowing down the analysis to groups of users, or even individual users, the \tool can log together with every request grouping information for that request. The simplest way of achieving this is to take advantage of the architecture of Flask applications in which a global \code{flask.request} is used to retrieve session information which can in turn is normally used to identify the user sending the request. From the user one can obtain the group.

\niceseparator

The following code snippet shows how we enabled \tool to enable user-by-user analysis\footnote{If we wanted to group by the user group for example, the code would change slightly by replacing ``user.id'' with ``user.group.id''}:

@@

In our case study, we wanted to understand the per-user performance of \epFeedItems 
--  an endpoint that retrieves a list of recommended articles for a given user.
It is the endpoint with the slowest response time and highest variability (see \Fref{fig:ep}).

% Once the \tool is configured with the line of code aforementioned.

@@

After configuring the \tool as shown earlier, the \perspective{Per-User Performance} perspective becomes available to  present the different response times for different users. Figure \ref{fig:tpu} presents a subset of the corresponding view in the \tool. The figure shows that the response times for this endpoint varied considerably for different users with some extreme cases where a user has to wait a full minute until their recommended articles were shown.

The reason for this behavior is that a user can be subscribed to anything from one to three dozen 
article sources and for each of the sources the system computed the personalized difficulty 
of each article at every request. Only after seeing the two perspectives presented in Figures \ref{fig:ep} and \ref{fig:tpu}, the maintainer refactored the architecture of the system to move the difficulty computation out of the interactive loop.
